---
title: "**Stock Module**"

author: |
  | **Bruno Caetano Vidigal**
  | Food and Agriculture Organization of the United Nations
output: pdf_document
---
---

```{r include=FALSE, cache=FALSE}
## Load required functions
library(data.table)
library(ggplot2)
library(scales)
library(knitr)
library(devtools)
#install_github("ndphillips/yarrr")
library("yarrr")
library(RColorBrewer)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE, cache=FALSE}

## This code was used to create the table tabGroup

# library(faosws)
# library(data.table)
# library(ggplot2)
# library(faoswsUtil)
# 
# ## set up for the test environment and parameters
# # initialYear = 1961 # Change documentation if this changes!
# 
# R_SWS_SHARE_PATH <- Sys.getenv("R_SWS_SHARE_PATH")
# 
# # This return FALSE if on the Statistical Working System
# if(CheckDebug()){
#   
#   message("Not on server, so setting up environment...")
#   
#   library(faoswsModules)
#   SETTINGS <- ReadSettings("modules/impute_stocks/sws.yml")
#   
#   # If you're not on the system, your settings will overwrite any others
#   R_SWS_SHARE_PATH <- SETTINGS[["share"]]
#   
#   # Define where your certificates are stored
#   SetClientFiles(SETTINGS[["certdir"]])
#   
#   # Get session information from SWS. Token must be obtained from web interface
#   GetTestEnvironment(baseUrl = SETTINGS[["server"]],
#                      token = SETTINGS[["token"]])
#   
# }
# 
# elements = c("51", "61", "71", "91")
# 
# # 51 production (Mt)
# # 61 imports (Mt)
# # 71 Variation Intial Exstenc (Mt) 
# # 91 exports (Mt)  
# # 101 feed (Mt)
# # 131 processed (Mt)
# # 141 food (Mt)
# 
# # Data
# key = DatasetKey(domain = "faostat_one", dataset = "FS1_SUA_UPD", dimensions = list(
#   Dimension(name = "geographicAreaFS", keys = GetCodeList("faostat_one", "FS1_SUA_UPD", "geographicAreaFS")[, code]),
#   Dimension(name = "measuredElementFS", keys = elements),
#   Dimension(name = "timePointYears", keys = as.character(2005:2012)),
#   Dimension(name = "measuredItemFS", keys = GetCodeList("faostat_one", "FS1_SUA_UPD", "measuredItemFS")[, code]))
# )
# data = GetData(key)
# data[, geographicAreaM49 := fs2m49(as.character(geographicAreaFS))]
# data[, measuredItemCPC := fcl2cpc(as.character(formatC(as.numeric(measuredItemFS), width = 4,
#                                                        flag = "0")))]
# # data = data[!is.na(geographicAreaM49) & 
# #               !(flagFaostat %in% c("M", "T"))]
# 
# data[, c("geographicAreaFS", "measuredItemFS", "flagFaostat") := NULL]
# setcolorder(data, c("geographicAreaM49", "measuredItemCPC", "measuredElementFS", 
#                     "timePointYears", "Value"))
# 
# # data[grepl("Hen", measuredItemCPC_description)]
# 
# data = data[, list(Value = sum(Value)), 
#             by = list(geographicAreaM49, measuredItemCPC, 
#                       measuredElementFS, timePointYears)]
# 
# item <- ReadDatatable("fbs_stocks_comm_codes")
# item <- item[fbs_key == "measuredItemCPC", fbs_code]
# 
# # Select just cereals, pulses and refined sugar
# 
# data <- data[measuredItemCPC %in% item]
# data[substr(measuredItemCPC, 1, 3) == "011", group := "cereals"]
# data[substr(measuredItemCPC, 1, 3) == "017", group := "pulses"]
# data[substr(measuredItemCPC, 1, 3) == "235", group := "refined_sugar"]
# data
# 
# ## More data manipulation 
# data <- dcast.data.table(data, geographicAreaM49 + measuredItemCPC + timePointYears + group ~ measuredElementFS,
#                          value.var = "Value")
# 
# setnames(data, old=c("51", "61", "71", "91"),
#          new=c("production", "imports", "variationIntialExstenc", "exports"))
# 
# ## Checking
# 
# for(cname in c("production", "imports", "variationIntialExstenc", "exports")){
#   data[is.na(get(cname)), c(cname) := 0]
# }
# 
# data[, netTrade := imports - exports]
# data
# 
# ## Country groups
# dlpath <- file.path("C:", "Users", "caetano", "Documents", "Github", "faoswsFeed",
#                     "data-raw", "IR_factor", "class.csv")
# 
# countryGroup <- fread(dlpath)
# countryGroup[, .N, GroupName]
# 
# countryGroup[grepl("income", GroupName), .N, GroupName]
# 
# countryIncomeGroup <- countryGroup[GroupCode %in% c("HIC", "LIC", "UMC", "LMC"), ]
# # countryRegionGroup <- countryGroup[GroupCode %in% c("NAC", "LCN", "AFR", "EAS", "EUU"),]
# 
# # library(countrycode)
# # ?countrycode
# 
# countryIncomeGroup[, geographicAreaM49 := as.character(countrycode(CountryCode, "wb", "iso3n"))]
# 
# # Sudan has the wrong name (it should be former Sudan)
# countryIncomeGroup[geographicAreaM49 == "736", CountryName := "Sudan (former)"]
# # China should be 1248
# countryIncomeGroup[geographicAreaM49 == "156", geographicAreaM49 := "1248"]
# #Exclude Channel Islands and Kosovo (not separately recognised by the UN)
# countryIncomeGroup <- countryIncomeGroup[!is.na(geographicAreaM49)]
# countryIncomeGroup
# 
# # Merge
# 
# data <- merge(data, countryIncomeGroup[, c("geographicAreaM49", "GroupName"), with = F], 
#       by = "geographicAreaM49", all.x=T)
# 
# data <- data[!is.na(geographicAreaM49)]
# data <- data[!is.na(GroupName)]
# 
# tabGroup <- data[, list(production = sum(production, na.rm = T),
#             imports = sum(imports, na.rm = T),
#             exports = sum(exports, na.rm = T),
#             deltaStocks = sum(variationIntialExstenc, na.rm = T)),
#      by=list(GroupName, group, timePointYears)]
# 
# tabGroup <- tabGroup[, trade := imports - exports]

tabGroup = fread("C:/Users/caetano/Documents/Github/faoswsStock/sandbox/tabGroup.csv")
```


## Abstract

This vignette provides analysis for Cereals, Pulses and Sugar.


# 1. Data

The data was pulled from the domain **faostat_one** and dataset **FS1_SUA_UPD** between **2005** and **2012** for all the countries.  
We are interested in **Cereals**, **Pulses** and **Sugar**.
The measuredElementFS pulled are:

```{r, echo=FALSE}
kable(data.table(measuredElementFS = c("51", "61", "71", "91"),
label = c("Production (Mt)", "Imports (Mt)", "Variation Intial Exstenc (Mt)", "Exports (Mt)")), format = "markdown", padding = 0)
```

We are using the .csv file from the **Feed** **Module** called **class**. This file has information for **income** by **country**. After combine the table from the **SWS** and the **class.csv** we get this dataset.
```{r, echo=FALSE}
head(tabGroup, 5)
```

# 2. Data Visualization

```{r, echo=FALSE}
options(warn=-1)

tabGroup$GroupName <- factor(tabGroup$GroupName, 
                        levels = c("High income", "Upper middle income", 
                                   "Lower middle income", "Low income"))
ggplot(data=tabGroup,
       aes(x=as.numeric(timePointYears), y=production, group=GroupName, col = GroupName)) +
  geom_line(aes(), stat = "identity", position=position_dodge(), size=1) +
  facet_wrap(~ group
             , scales = "free"
  ) +
  theme(legend.title=element_blank(), strip.text = element_text(size=7)) +
  xlab('Year') + ylab('Production') +
  scale_x_continuous(lim=c(2005, 2012), breaks=seq(2005, 2012, 1)) +
 theme(
    axis.text.x = element_text(size = 8, face = "bold", angle = 45, vjust = .5),
    legend.position = "top"
  )
```

```{r, echo=FALSE}
options(warn=-1)
ggplot(data=tabGroup,
       aes(x=as.numeric(timePointYears), y=trade, group=GroupName, col = GroupName)) +
  geom_line(aes(), stat = "identity", position=position_dodge(), size=1) +
  facet_wrap(~ group
             , scales = "free"
  ) +
  theme(legend.title=element_blank(), strip.text = element_text(size=7)) +
  xlab('Year') + ylab('Imports - Exports') +
  scale_x_continuous(lim=c(2005, 2012), breaks=seq(2005, 2012, 1)) +
  geom_hline(yintercept=0) +
   theme(
    axis.text.x = element_text(size = 8, face = "bold", angle = 45, vjust = .5),
    legend.position = "top"
  )
```
