---
title: "**Stock Module**"

author: |
  | **Bruno Caetano Vidigal**
  | Food and Agriculture Organization of the United Nations
output: pdf_document
---
---

```{r include=FALSE, cache=FALSE}
## create R program in working directory
## knitr::purl('sandbox/amis/descriptionAmis.Rmd')
## Load required functions
library(data.table)
library(ggplot2)
library(scales)
library(knitr)
library(devtools)
library("yarrr")
library(RColorBrewer)
library(faoswsUtil)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE, cache=FALSE}
dlpath <- file.path("C:", "Users", "caetano", "Documents", "Github", "faoswsStock",
                    "sandbox", "validation", "2016_09_15_final_result_stocks.csv")
dataValidation <- fread(dlpath)
```


## Abstract

This vignette provides a detailed analysis of the stock module. In addition to analysing the delta stocks, this document has analysis
about the **opening** **stocks** and **%** **of** **stocks** available.  
The group of commodities studied are **cereals**, **pulses** and **refined** **sugar**. 

# 1. Descriptive Statistics

The methodology used to imputate the delta stock data was based on previous analyses done with **AMIS** **Data** and **F**.**O** **Lichts** **Data**. 
Find below the basic description of the data.

```{r, echo=FALSE}
basic_info <- dataValidation[, list(countries = length(unique(geographicAreaM49)),
commodities = length(unique(measuredItemCPC)),
min_year = min(timePointYears),
max_year = max(timePointYears))]

kable(basic_info, format = "markdown", padding = 0, 
col.names=c("Country", "Commodity", "Min Year", "Max Year"))
```

# 2. Data Visualization

There is no production for **refined** **sugar** in **2014**.

## 2.1 World


```{r, echo=FALSE, fig.height=6}
options(warn=-1)

dataValidation[, timePointYears := as.numeric(timePointYears)]

tabPercent <- dataValidation[, list(
          percent = sum(openingStocks, na.rm=T)/sum(totalSupply, na.rm = T)
), by = list(timePointYears, itemGroup)
]

plotcolors <- colorRampPalette(c("black", "blue", "green"))

ggplot(data=tabPercent,
      aes(x=timePointYears, y=percent, colour=itemGroup)) +
 geom_line(aes(), stat = "identity", position=position_dodge(), size=1) + 
scale_y_continuous(labels = scales::percent) +
  scale_color_manual( values = plotcolors(length(unique(tabPercent$itemGroup))) ) +
  theme(legend.title=element_blank(), strip.text = element_text(size=7)) +
  ylab('% Opening Stocks') + xlab('') +
  ggtitle("World: % Opening Stocks by Year") +
  scale_x_continuous(lim=c(1999, 2014), breaks=seq(1999, 2014, 3)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 8, face = "bold", angle = 45, vjust = .5),
    legend.position = "top",
    legend.title=element_blank()
)
  
```

```{r, echo=FALSE, fig.height=6}
options(warn=-1)

tabSupply <- dataValidation[, list(
totalSupply = sum(totalSupply, na.rm = T)
), by = list(timePointYears, itemGroup)
]

plotcolors <- colorRampPalette(c("black", "blue", "green"))

ggplot(data=tabSupply,
      aes(x=timePointYears, y=totalSupply, colour=itemGroup)) +
 geom_line(aes(), stat = "identity", position=position_dodge(), size=1) + 
 facet_grid(
    itemGroup ~ .
      , scales = "free"
    ) +
# scale_y_continuous(labels = scales::percent) +
  scale_color_manual( values = plotcolors(length(unique(tabSupply$itemGroup))) ) +
  theme(legend.title=element_blank(), strip.text = element_text(size=7)) +
  ylab('Total Supply (tonnes)') + xlab('') +
  ggtitle("World: Total Supply by Year") +
  scale_x_continuous(lim=c(1999, 2014), breaks=seq(1999, 2014, 3)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 8, face = "bold", angle = 45, vjust = .5),
    legend.position = "top",
    legend.title=element_blank()
)
  
```

```{r, echo=FALSE, fig.height=6}
options(warn=-1)

tabDelta <- dataValidation[, list(
totalDeltaStocks = sum(deltaStocks, na.rm = T),
totalSupply = sum(totalSupply, na.rm = T)
), by = list(timePointYears, itemGroup)
]

setkey(tabDelta, itemGroup, timePointYears)
tabDelta[, deltaTotalSupply := c(0, diff(totalSupply)),
         by = list(itemGroup)]


ggplot(data = tabDelta[timePointYears > 1999 & timePointYears < 2014],
       aes(x=deltaTotalSupply, y=totalDeltaStocks, group=1)) +
  geom_point(aes(), stat = "identity", position=position_dodge(), size=1.5) +
  facet_wrap(~ itemGroup, scales = "free") +
  theme_bw() +
  theme(legend.title=element_blank(), strip.text = element_text(size=7)) +
  ggtitle("World \n Delta Total Supply vs. Delta Stocks") +
  ylab('Delta Stocks (Million tonnes)') + xlab('Delta Total Supply (Million tonnes)') +
  geom_smooth(method=lm)

```

## 2.2 Income Group

```{r, echo=FALSE, fig.height=6}
options(warn=-1)

tabPercent <- dataValidation[, list(
          percent = sum(openingStocks, na.rm=T)/sum(totalSupply, na.rm = T)
# ,
# totalSupply = sum(totalSupply, na.rm = T)
), by = list(timePointYears, itemGroup, incomeGroup)
]

tabPercent$incomeGroup <- factor(tabPercent$incomeGroup, 
levels = c("High income", "Upper middle income", "Lower middle income", "Low income"))

plotcolors <- colorRampPalette(c("gray70", "blue", "green", "black"))
tabPercent <- tabPercent[!is.na(incomeGroup)]

ggplot(data=tabPercent,
      aes(x=timePointYears, y=percent, group=incomeGroup, colour = incomeGroup)) +
 geom_line(aes(), stat = "identity", position=position_dodge(), size=1) + 
 facet_grid(
    itemGroup ~ .
      , scales = "free"
    ) +
scale_y_continuous(labels = scales::percent) +
   scale_color_manual( values = plotcolors(length(unique(tabPercent$incomeGroup)))) +
   theme(legend.title=element_blank(), strip.text = element_text(size=5)) +
  ylab('% Opening Stocks') + xlab('') +
  ggtitle("Income Group: % Opening Stocks by Year") +
  scale_x_continuous(lim=c(1999, 2014), breaks=seq(1999, 2014, 3)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 8, face = "bold", angle = 45, vjust = .5),
    legend.position = "top",
    legend.title=element_blank()
)

````

```{r, echo=FALSE, fig.height=6}
options(warn=-1)

tabSupplyIncome <- dataValidation[, list(
totalSupply = sum(totalSupply, na.rm = T)
), by = list(timePointYears, itemGroup, incomeGroup)
]

tabSupplyIncome$incomeGroup <- factor(tabSupplyIncome$incomeGroup, 
levels = c("High income", "Upper middle income", "Lower middle income", "Low income"))

plotcolors <- colorRampPalette(c("gray70", "blue", "green", "black"))
tabSupplyIncome <- tabSupplyIncome[!is.na(incomeGroup)]

ggplot(data=tabSupplyIncome,
      aes(x=timePointYears, y=totalSupply, group=incomeGroup, colour = incomeGroup)) +
 geom_line(aes(), stat = "identity", position=position_dodge(), size=1) + 
 facet_grid(
    itemGroup ~ .
      , scales = "free"
    ) +
# scale_y_continuous(labels = scales::percent) +
   scale_color_manual( values = plotcolors(length(unique(tabSupplyIncome$incomeGroup)))) +
   theme(legend.title=element_blank(), strip.text = element_text(size=5)) +
  ylab('Total Supply (tonnes)') + xlab('') +
  ggtitle("Income Group: Total Supply by Year") +
  scale_x_continuous(lim=c(1999, 2014), breaks=seq(1999, 2014, 3)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 8, face = "bold", angle = 45, vjust = .5),
    legend.position = "top",
    legend.title=element_blank()
)

````

```{r, echo=FALSE, fig.height=8, fig_width = 4}
options(warn=-1)

tabDeltaIncome <- dataValidation[, list(
totalDeltaStocks = sum(deltaStocks, na.rm = T),
totalSupply = sum(totalSupply, na.rm = T)
), by = list(timePointYears, itemGroup, incomeGroup)
]


tabDeltaIncome$incomeGroup <- factor(tabDeltaIncome$incomeGroup, 
levels = c("High income", "Upper middle income", "Lower middle income", "Low income"))

plotcolors <- colorRampPalette(c("gray70", "blue", "green", "black"))
tabDeltaIncome <- tabDeltaIncome[!is.na(incomeGroup)]

setkey(tabDeltaIncome, itemGroup, incomeGroup, timePointYears)
tabDeltaIncome[, deltaTotalSupply := c(0, diff(totalSupply)),
         by = list(itemGroup, incomeGroup)]


ggplot(data = tabDeltaIncome[timePointYears > 1999 & timePointYears < 2014 & itemGroup == "cereals"],
       aes(x=deltaTotalSupply, y=totalDeltaStocks, group=incomeGroup, colour=incomeGroup)) +
  geom_point(aes(), stat = "identity", position=position_dodge(), size=1.5) +
  facet_grid(itemGroup ~ incomeGroup, scales="free_y") +
  # facet_grid(incomeGroup ~ type, scales="free_y") +
  scale_color_manual( values = plotcolors(length(unique(tabDeltaIncome$incomeGroup)))) +
  theme_bw() +
  theme(legend.title=element_blank(), strip.text = element_text(size=7)) +
  ggtitle("Income Group & Cereals \n Delta Total Supply vs. Delta Stocks") +
  ylab('Delta Stocks (Million tonnes)') + xlab('Delta Total Supply (Million tonnes)') #+
  #geom_smooth(method=lm)
  
```

```{r, echo=FALSE, fig.height=6}
options(warn=-1)

ggplot(data = tabDeltaIncome[timePointYears > 1999 & timePointYears < 2014 & itemGroup == "pulses"],
       aes(x=deltaTotalSupply, y=totalDeltaStocks, group=incomeGroup, colour=incomeGroup)) +
  geom_point(aes(), stat = "identity", position=position_dodge(), size=1.5) +
  facet_grid(itemGroup ~ incomeGroup, scales="free_y") +
  scale_color_manual( values = plotcolors(length(unique(tabDeltaIncome$incomeGroup)))) +
  theme_bw() +
  theme(legend.title=element_blank(), strip.text = element_text(size=7)) +
  ggtitle("Income Group & Pulses \n Delta Total Supply vs. Delta Stocks") +
  ylab('Delta Stocks (Million tonnes)') + xlab('Delta Total Supply (Million tonnes)') #+
  #geom_smooth(method=lm)

````

```{r, echo=FALSE, fig.height=6}
options(warn=-1)

ggplot(data = tabDeltaIncome[timePointYears > 1999 & timePointYears < 2014 & itemGroup == "refined_sugar"],
       aes(x=deltaTotalSupply, y=totalDeltaStocks, group=incomeGroup, colour=incomeGroup)) +
  geom_point(aes(), stat = "identity", position=position_dodge(), size=1.5) +
  facet_grid(itemGroup ~ incomeGroup, scales="free_y") +
  scale_color_manual( values = plotcolors(length(unique(tabDeltaIncome$incomeGroup)))) +
  theme_bw() +
  theme(legend.title=element_blank(), strip.text = element_text(size=7)) +
  ggtitle("Income Group & Refined Sugar \n Delta Total Supply vs. Delta Stocks") +
  ylab('Delta Stocks (Million tonnes)') + xlab('Delta Total Supply (Million tonnes)') #+
  #geom_smooth(method=lm)

````


## 2.3 Country

```{r, echo=FALSE, fig.height=6}
options(warn=-1)

tabPercentCountry <- dataValidation[, list(
          percent = sum(openingStocks, na.rm=T)/sum(totalSupply, na.rm = T)
# ,
# totalSupply = sum(totalSupply, na.rm = T)
), by = list(timePointYears, itemGroup, geographicAreaM49_description)
]

tabPercentCountry <- tabPercentCountry[!is.na(percent)]

plotcolors <- colorRampPalette(c("gray70", "blue", "green", "black"))
# plotcolors <- colorRampPalette(c("gray70", "blue", "black"))

ggplot(data=tabPercentCountry[geographicAreaM49_description %in% c("Brazil", "India", "Germany", "United States of America")],
      # aes(x=timePointYears, y=percent, colour=type)) +
      aes(x=timePointYears, y=percent, group=geographicAreaM49_description, colour = geographicAreaM49_description)) +
 geom_line(aes(), stat = "identity", position=position_dodge(), size=1) + 
 facet_grid(
    itemGroup ~ .
      , scales = "free"
    ) +
scale_y_continuous(labels = scales::percent) +
    scale_color_manual( values = plotcolors(length(unique(tabPercentCountry[geographicAreaM49_description %in% c("Brazil", "India", "Germany", "United States of America", "China")]$geographicAreaM49_description)))) +
   theme(legend.title=element_blank(), strip.text = element_text(size=5)) +
  ylab('% Opening Stocks') + xlab('') +
  ggtitle("Country: % Opening Stocks by Year") +
  scale_x_continuous(lim=c(1999, 2014), breaks=seq(1999, 2014, 3)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 8, face = "bold", angle = 45, vjust = .5),
    legend.position = "top",
    legend.title=element_blank()
)
````

```{r, echo=FALSE, fig.height=6}
options(warn=-1)

tabSupplyCountry <- dataValidation[, list(
totalSupply = sum(totalSupply, na.rm = T)
), by = list(timePointYears, itemGroup, geographicAreaM49_description)
]

tabSupplyCountry <- tabSupplyCountry[!is.na(totalSupply)]

plotcolors <- colorRampPalette(c("gray70", "blue", "green", "black"))
# plotcolors <- colorRampPalette(c("gray70", "blue", "black"))

ggplot(data=tabSupplyCountry[geographicAreaM49_description %in% c("Brazil", "India", "Germany", "United States of America")],
      aes(x=timePointYears, y=totalSupply, group=geographicAreaM49_description, colour = geographicAreaM49_description)) +
 geom_line(aes(), stat = "identity", position=position_dodge(), size=1) + 
 facet_grid(
    itemGroup ~ .
      , scales = "free"
    ) +
# scale_y_continuous(labels = scales::percent) +
    scale_color_manual( values = plotcolors(length(unique(tabSupplyCountry[geographicAreaM49_description %in% c("Brazil", "India", "Germany", "United States of America")]$geographicAreaM49_description)))) +
   theme(legend.title=element_blank(), strip.text = element_text(size=5)) +
  ylab('Total Supply (tonnes)') + xlab('') +
  ggtitle("Country: Total Supply by Year") +
  scale_x_continuous(lim=c(1999, 2014), breaks=seq(1999, 2014, 3)) +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 8, face = "bold", angle = 45, vjust = .5),
    legend.position = "top",
    legend.title=element_blank()
)
````

```{r, echo=FALSE, fig.height=8, fig_width = 4}
options(warn=-1)

tabDeltaCountry <- dataValidation[, list(
totalDeltaStocks = sum(deltaStocks, na.rm = T),
totalSupply = sum(totalSupply, na.rm = T)
), by = list(timePointYears, itemGroup, geographicAreaM49_description)
]

tabDeltaCountry <- tabDeltaCountry[!is.na(totalSupply)]

plotcolors <- colorRampPalette(c("gray70", "blue", "green", "black"))

setkey(tabDeltaCountry, itemGroup, geographicAreaM49_description, timePointYears)
tabDeltaCountry[, deltaTotalSupply := c(0, diff(totalSupply)),
         by = list(itemGroup, geographicAreaM49_description)]


ggplot(data = tabDeltaCountry[timePointYears > 1999 & timePointYears < 2014 &
geographicAreaM49_description %in% c("Brazil", "India", "Germany", "United States of America") & itemGroup == "cereals"],
       aes(x=deltaTotalSupply, y=totalDeltaStocks, group=geographicAreaM49_description, colour=geographicAreaM49_description)) +
  geom_point(aes(), stat = "identity", position=position_dodge(), size=1.5) +
  facet_grid(itemGroup ~ geographicAreaM49_description, scales="free_y") +
  scale_color_manual( values = plotcolors(length(unique(tabSupplyCountry[geographicAreaM49_description %in% c("Brazil", "India", "Germany", "United States of America")]$geographicAreaM49_description)))) +
  theme_bw() +
  theme(legend.title=element_blank(), strip.text = element_text(size=7)) +
  ggtitle("Country & Cereals \n Delta Total Supply vs. Delta Stocks") +
  ylab('Delta Stocks (Million tonnes)') + xlab('Delta Total Supply (Million tonnes)') #+
  #geom_smooth(method=lm)
  
```

```{r, echo=FALSE, fig.height=8, fig_width = 4}
options(warn=-1)

ggplot(data = tabDeltaCountry[timePointYears > 1999 & timePointYears < 2014 &
geographicAreaM49_description %in% c("Brazil", "India", "Germany", "United States of America") & itemGroup == "pulses"],
       aes(x=deltaTotalSupply, y=totalDeltaStocks, group=geographicAreaM49_description, colour=geographicAreaM49_description)) +
  geom_point(aes(), stat = "identity", position=position_dodge(), size=1.5) +
  facet_grid(itemGroup ~ geographicAreaM49_description, scales="free_y") +
  scale_color_manual( values = plotcolors(length(unique(tabSupplyCountry[geographicAreaM49_description %in% c("Brazil", "India", "Germany", "United States of America")]$geographicAreaM49_description)))) +
  theme_bw() +
  theme(legend.title=element_blank(), strip.text = element_text(size=7)) +
  ggtitle("Country & Pulses \n Delta Total Supply vs. Delta Stocks") +
  ylab('Delta Stocks (Million tonnes)') + xlab('Delta Total Supply (Million tonnes)') #+
  #geom_smooth(method=lm)

````

```{r, echo=FALSE, fig.height=8, fig_width = 4}
options(warn=-1)

ggplot(data = tabDeltaCountry[timePointYears > 1999 & timePointYears < 2014 &
geographicAreaM49_description %in% c("Brazil", "India", "Germany", "United States of America") & itemGroup == "refined_sugar"],
       aes(x=deltaTotalSupply, y=totalDeltaStocks, group=geographicAreaM49_description, colour=geographicAreaM49_description)) +
  geom_point(aes(), stat = "identity", position=position_dodge(), size=1.5) +
  facet_grid(itemGroup ~ geographicAreaM49_description, scales="free_y") +
  scale_color_manual( values = plotcolors(length(unique(tabSupplyCountry[geographicAreaM49_description %in% c("Brazil", "India", "Germany", "United States of America")]$geographicAreaM49_description)))) +
  theme_bw() +
  theme(legend.title=element_blank(), strip.text = element_text(size=7)) +
  ggtitle("Country & Refined Sugar \n Delta Total Supply vs. Delta Stocks") +
  ylab('Delta Stocks (Million tonnes)') + xlab('Delta Total Supply (Million tonnes)') #+
  #geom_smooth(method=lm)
````
